{\fontfamily{ptm}\selectfont Inexact Accelerated Proximal Gradient }

### 2025-10-02 Notebook LaTEX/Inexact AGP made simple for applications/main.tex
>w<, changes

\title{{\fontfamily{ptm}\selectfont Inexect Accelerated Proximal Gradient }}

@@

@@

            Let $(f, g, L)$ satisfies Assumption \ref{ass:for-inxt-pg-ineq}, $\epsilon \ge 0, \rho > 0$, then for all $x \approx_\epsilon T_\rho(x)$, it has the following equivalent representations: 

@@

@@

@@

        In this section we will show an optimization problem that allows us to solve for some $\tilde x \approx_\epsilon \hprox_{\lambda g}(z)$. 
        Eventually we want to evaluate $T_{\rho}(x)$ of some $F = f + g$ inexactly and, by Lemma \ref{lemma:other-repr-inxt-pg}, one can achieve that through inexact evaluation of $\hprox_{\rho^{-1}g}$ in the sense of Definition \ref{def:inxt-pp}. 
        Most of these results are from the literature. 

@@

        \begin{theorem}[primal translate to dual]\label{thm:primal-dual-trans}

@@

        This theorem that follows is from Villa et al. \cite[Proposition 2.3]{villa_accelerated_2013}, but put into our symbols and, Definition 
        \begin{theorem}[duality gap of inexact proximal problem]\label{thm:dlty-gap-inxt-pp}

@@

        The following fact is Proposition \cite[Theorem 5.1]{villa_accelerated_2013}. 
        \begin{fact}[minimizing dual of the proximal problem]\label{fact:minimizing-dual-pp}
            Let $\bar v$ be a solution of $\Psi$. 
            Suppose that $(v_n)_{n \ge 0}$ is a minimizing sequence for $\Psi$. 

@@

        We remark that the above fact translates any algorithm that optimizes the function value of the dual problem into optimizing duality gap $\mathbf G(z_n, v_n)$. 
        For this reason, the number of iterations of the inner loop required to achieve $\mathbf G(z_n, v_n) < \epsilon$ for a given $e$ is related to the convergence rate of the algorithms used to optimize $\Psi_\lambda(v_n)$. 

@@

\section{The accelerated proximal gradient with controlled errors}

@@

            \item $(B_k)_{k \ge 0}$ is a non-negative sequence, characterizing the potential line search routine. 

@@

        Let $(f, g, L)$, and $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}$, be given by Definition \ref{def:inxt-apg}. 

@@

    \begin{proposition}[valid error schedule and convergence rate]\;\label{prop:inxt-apg-cnvg-generic}\\
        Let $(f, g, L)$, $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}$ be given by Definition \ref{def:inxt-apg}. 
        Denote for brevity $\beta_0 = 1$, $\beta_k = \prod_{i = 1}^{k} \max\left(1 - \alpha_i, \frac{\alpha_i^2L_i}{\alpha_{i - 1}^2L_{i - 1}}\right)$ and $L_k = B_k + \rho_k$. 
        If for some fixed $\mathcal E_0 \ge 0, p \ge 1$ the parameter $\rho_k, \epsilon_k$ can satisfy for all $k \ge 0$ the condition

@@

        Then for the sequence generated $(y_k, x_k, v_k)_{k \ge 0}$ by the algorithm, for all $k \ge 0$ they satisfy: 

@@

        These conditions are equivalent to the assumption that $\frac{- \mathcal E_0\beta_k}{k^p} \le \frac{\rho_k}{2}\Vert x_k - y_k\Vert^2 - \epsilon_k$. 

@@

\section{Linear convergence for the inner loop proximal problem}

@@

    To attain $x_k$, one approach is to utilise Theorem \ref{thm:dlty-gap-inxt-pp} numerically, usually using iterative algorithms. 

@@

            \item Let $\tau > 0$ be the step size inverse, let $T_{\tau}$ be the proximal operator of $f + g$ as given by $T_{\tau}(x) := \hprox_{\tau^{-1}g}(x - \tau^{-1}\nabla f(x))$, 
            \item the gradient mapping $\mathcal G_{\tau}$ be given by: $\mathcal G_t(x) = \tau (x - T_t(x))$, 

@@

@@

            \item $(g, \omega, A)$ satisfies Assumption \ref{ass:for-inxt-prox}. 
            \item In addition, function $\Psi_\lambda$ as given by \eqref{eqn:primal-pp} satifies gradient mapping error bound (Assumption \ref{ass:pg-eb}) where, $f(v) = \frac{1}{2\lambda}\Vert \lambda A^\top v - y\Vert^2, g(v) = \omega(A v) - \frac{1}{2\lambda}\Vert y\Vert^2$. 

@@

                F(z) - F(\tilde x) + \frac{\tau}{2}\Vert x - z\Vert^2 - \frac{\tau}{2}\Vert z - \tilde x\Vert^2

@@

                F(z) - F(\tilde x) - \frac{\tau}{2}\Vert x - \tilde x\Vert^2

@@

                &= F(z) - F(\tilde x) - \frac{1}{2\tau}\Vert \mathcal G_\tau(x) \Vert^2

@@

                &\le F(z) - F(\tilde x) - \frac{1}{2\tau}\Vert \mathcal G_\tau(x)\Vert^2 

@@

                + \Vert \mathcal G_\tau(x)\Vert^2\left(

@@

            Now, for all $z = \bar x \in S$, from Assumption \ref{ass:eb-for-pp} it has

@@

    \subsection{characterizing linear convergence of the proximal problem}

>>>>

\title{{\fontfamily{ptm}\selectfont Inexact Accelerated Proximal Gradient }}

@@

        Let $\rho > 0$, the exact proximal gradient operator defined for $(f, g, L)$ satisfying Assumption \ref{ass:for-inxt-pg-ineq} has
            T_{\rho}(x) &= \argmin_{z\in\RR^n}\left\lbrace g(z) + \langle \nabla f(x), z\rangle + \frac{\rho}{2}\Vert z - x\Vert^2 \right\rbrace
            &= \hprox_{\rho^{-1} g}\left(x - \rho^{-1}\nabla f(x)\right). 
        The following definition extends the proximal gradient operator to the inexact case using the concept of $\epsilon$-subgradient as given by Definition \ref{def:esp-subgrad}. 

@@

            Let $(f, g, L)$ satisfies Assumption \ref{ass:for-inxt-pg-ineq}, $\epsilon \ge 0, \rho > 0$, then for all $\tilde x \approx_\epsilon T_\rho(x)$, it has the following equivalent representations: 

@@

            We also used $F = f + g$. 

@@

            In this inequality, observe that the parameter $\epsilon$ controls the inexactiness of the proximal gradient evaluation. 
            Morespecifically, $\epsilon_k$ controls the absolute pertubatnions of the proximal gradient inequality compared to its exact counterpart. 
            $\rho$ on the otherhand, it is the over-relaxation of proximal gradient operator and it compensates the pertubations caused by $\epsilon$ relative to the term $\Vert \tilde x - x\Vert^2$. 

@@

        In this section we will present the optimization problem that obtains a $\tilde x$ such that $\tilde x \approx_\epsilon \hprox_{\lambda g}(z)$. 
        Eventually we want to evaluate $T_{\rho}(x)$ of some $F = f + g$ inexactly using Lemma \ref{lemma:other-repr-inxt-pg}. 
        To do that one would need to evaluate $\hprox_{\rho^{-1}g}$ inexactly which is defined in Definition \ref{def:inxt-pp}. 
        Most of these results that will follow are from the literature. 

@@

        \begin{theorem}[primal translate to dual {\cite[Proposition 2.2]{villa_accelerated_2013}}]\label{thm:primal-dual-trans}

@@

        This theorem that follows is from Villa et al. \cite[Proposition 2.3]{villa_accelerated_2013}. 
        \begin{theorem}[duality gap of inexact proximal problem {\cite[Proposition 2.3]{villa_accelerated_2013}}]\;\label{thm:dlty-gap-inxt-pp}\\

@@

        The following fact is Proposition is from Villa et al. \cite[Theorem 5.1]{villa_accelerated_2013}. 
        \begin{fact}[minimizing dual of the proximal problem {\cite[Theorem 5.1]{villa_accelerated_2013}}]\label{fact:minimizing-dual-pp}
            Let $\bar v$ be a solution of $\Psi_\lambda$. 
            Suppose that $(v_n)_{n \ge 0}$ is a minimizing sequence for $\Psi_\lambda$. 

@@

        We remark that the above fact translates any algorithm that optimizes the function value of the dual problem $\Psi_\lambda$ into optimizing duality gap $\mathbf G(z_n, v_n)$. 
        For this reason, the number of iterations of the inner loop required to achieve $\mathbf G(z_n, v_n) < \epsilon$ for a given $\epsilon$ is related to the convergence rate of the algorithms used to optimize $\Psi_\lambda(v_n)$. 
        With the theorem derived above, and using Theorem \ref{thm:dlty-gap-inxt-pp} it implies that any algorithm which can optimize function value $\Psi_\lambda$ will produce iterates sufficient to achieve $\approx_\epsilon \hprox_{\lambda g}(y)$. 

@@

\section{The inexact accelerated proximal gradient with controlled errors}

@@

            \item $(B_k)_{k \ge 0}$ is a non-negative sequence, characterizing any potential line search routine. 

@@

        Let $(F, f, g, L)$, and $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}$, be given by Definition \ref{def:inxt-apg}. 

@@

    The following assumption encapsulate assumptions on the errors such that a near optimal convergence rate is still attainable by an algorithm that satisfies Definition \ref{def:inxt-apg}. 
    \begin{assumption}[valid error schedule]\label{ass:valid-err-schedule}
        The following assumption is about an algorithm satisfying Definition \ref{def:inxt-apg}, its parameters $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}$ in relation to its iterates $(y_k, x_k, v_k)_{k\ge 0}$ and, some additional parameters $(\beta_k)_{k\ge 0}, \mathcal E_0$ and $p$. 
        Let 
            \item $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}, (F, f, g, L)$ and $(y_k, x_k, v_k)_{k\ge 0}$ be given by Definition \ref{def:inxt-apg}. 
            \item $\mathcal E_0 \ge 0$ be arbitrary;
            \item the sequence $(\beta_k)_{k\ge 0}$ be defined as $\beta_k := \prod_{i = 1}^{k} \max\left(1 - \alpha_i, \frac{\alpha_i^2L_i}{\alpha_{i - 1}^2L_{i - 1}}\right)$ for all $k \ge 1$, with the base case being $\beta_0 = 1$; 
            \item $p \ge 1$ is some constant which will bound the error $\epsilon_k$ relative to $\rho_k$. 
        In addition, we assume that the error parameter $\epsilon_k$ and over-relaxation parameter $\rho_k$, iterates $x_k, y_k$ and $\beta_k$ together satisfies for all $k \ge 0$ the relations:

@@

    \begin{proposition}[generic convergence rate under valid error schedule]\;\label{prop:inxt-apg-cnvg-generic}\\
        Let $(F, f, g, L)$, $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}$, $(\beta_k)_{k\ge 0}, \mathcal E_0, p$ as assumed in Assumption \ref{ass:valid-err-schedule}. 
        Then for the iterates generated $(y_k, x_k, v_k)_{k \ge 0}$ by the algorithm, for all $k \ge 0$ they will satisfy: 

@@

        These conditions are equivalent to the assumption that $\frac{- \mathcal E_0\beta_k}{k^p} \le \frac{\rho_k}{2}\Vert x_k - y_k\Vert^2 - \epsilon_k$ (which was stated in Assumption \ref{ass:valid-err-schedule}). 

@@

    \subsection{convergence results of the outer loop}
        This section will give specific instances of the error control sequence $(\epsilon_k)_{k \ge 0}, (\rho_k)_{k \ge0}$ and, momentum sequence $(\alpha_k)_{k \ge0}$ such that an optimal convergence rate of $\mathcal O(1/k^2)$ can be achieved. 
        \begin{proposition}[$\mathcal O(1/k^2)$ optimal convergence rate of the outer loop]\;\label{prop:opt-cnvg-outr-loop}\\
            Let $(f, g, L)$, $(\alpha_k, B_k, \rho_k, \epsilon_k)_{k \ge 0}$, $(\beta_k)_{k\ge 0}, \mathcal E_0, p$ as assumed in Assumption \ref{ass:valid-err-schedule}. 
            Assume in addition that
                \item there exists $\bar x \in \RR^n$ that is an minimizer of $F = f + g$;
                \item $\alpha_0 = 1, p > 0$; 
                \item the sequence $(\alpha_k)_{k\ge 0}$ satisfies for all $k \ge 0$ the equality $(1 - \alpha_k) = \alpha_{k}^2L_k\alpha_{k - 1}^{-2}L_{k - 1}^{-1}$;
                \item the sequence $L_k := B_k + \rho_k$ is bounded, and there exists an $L_{\max}$ such that for all $k \ge 0$ it has $L_{\max} \ge \max_{k\ge i\ge 0} L_i$. 
            Then, $\alpha_k \in (0, 1]$ hence valid and, it satisfies for all $k \ge 0$: 
                    1 + \frac{k\alpha_0\sqrt{L_0}}{2\sqrt{L_{\max}}}
                \right)^{-2}\left(
            Finally, since $p > 0$ the sum is convergent and hence it has an over all convergence rate $\mathcal O(1/k^2)$. 
            From the assumption that $(\alpha_k)_{k \ge 0}$ has $(1 - \alpha_k) = \alpha_k^2L_k \alpha_0^{-2}L_0^{-1}$ for all $k \ge 0$ and, the definition of $\beta_k$, it yields the following equalities: 
                \beta_k = \prod_{i = 1}^k \max\left(
                    1 - \alpha_i, \frac{\alpha_i^2L}{\alpha_{i - 1}^2L_{i - 1}}
                = \prod_{i = 1}^k(1 - \alpha_i) 
                = \prod_{i = 1}^k \frac{\alpha_i^2L}{\alpha_0^2L_0} = \frac{\alpha_k^{2}L_k}{\alpha_0^2L_0}. 
            With the above relation, and the definitions of the sequences $(\alpha_k)_{k \ge 0}, (\beta_k)_{k\ge 0}$ it satisfies for all $k \ge 1$ the properties: 
                \item $\beta_k$ is nonincreasing and $\beta_k > 0$ for all $k \ge 0$ because $\beta_k = \prod_{i = 1}^{k} (1 - \alpha_i)$ and, $\alpha_k \in (0, 1]$. 
                \item It has the equalities $\beta_k/\beta_{k - 1} = (1 - \alpha_k) = \frac{\alpha_k^2L_k}{\alpha_{k - 1}^2 L_{k - 1}}$ for all $k \ge 1$. 
            Using the above observations, we can show the chain of equalities $\alpha_k^{2} = (1 - \beta_k/\beta_{k - 1})^2 = \beta_kL_0\alpha_0^2L_k^{-1}$ for all $k \ge 0$. 
            This is true by first considering the relations $\prod_{i = 1}^k(1 - \alpha_i) = \beta_k$: 
            \begin{align}\label{eqn:opt-cnvg-outr-loop-pitem1}\begin{split}
                (1 - \alpha_k) &= \beta_k/\beta_{k - 1}
                \alpha_k &= 1 - \beta_k / \beta_{k - 1}
                \implies 
                \alpha_k^2 &= (1 - \beta_k / \beta_{k - 1})^2.     
            Next, the recursive relation of $(\alpha_k)_{k \ge 0}$ gives
            \begin{align}\label{eqn:opt-cnvg-outr-loop-pitem2}\begin{split}
                \alpha_k^2&= (1 - \alpha_k)\alpha_{k - 1}^2L_{k - 1}L_k^{-1}
                (1 - \alpha_k)
                \frac{\alpha_{k - 1}^2L_{k - 1}}{\alpha_0^2L_0}
                \frac{L_{k - 1}\alpha_0^2L_0}{L_k}
                (\beta_k\beta_{k - 1}^{-1})\beta_{k - 1}L_0\alpha_0^2L_k^{-1}
                &= \beta_kL_0\alpha_0^2L_k^{-1}.         
            Combining \eqref{eqn:opt-cnvg-outr-loop-pitem1}, \eqref{eqn:opt-cnvg-outr-loop-pitem2} it would mean for all $i \ge 1$ it has: 
                L_0 \alpha_0^2 L_i^{-1} &= 
                \beta_i^{-1}\left(
                    1 - \frac{\beta_k}{\beta_{k - 1}}
                \right)^2
                \beta_i \left(
                    \beta_i^{-1} - \beta_{i - 1}^{-1}
                \right)^2
                \beta_i \left(
                    \beta_i^{-1/2} - \beta_{i - 1}^{-1/2}
                \right)^2
                    \beta_i^{-1/2} + \beta_{i - 1}^{-1/2}
                \right)^2
                \beta_i \left(
                    \beta_i^{-1/2} - \beta_{i - 1}^{-1/2}
                \right)^2
                    2\beta_i^{-1/2}
                \right)^2
                &= 4\left(
                    \beta_i^{-1/2} - \beta_{i - 1}^{-1/2}
                \right)^2
                \implies
                \sqrt{\frac{L_0\alpha_0^2}{L_i}} &\le
                2 \left(
                    \beta_i^{-1/2} - \beta_{i - 1}^{-1/2}
            Telescope for all $i \ge 1$: 
                \frac{\alpha_0\sqrt{L_0}}{2}\sum_{i = 1}^{k}\sqrt{L^{-1}_k}
                \beta_k^{-1/2} - \beta_{0}^{-1/2} = 
                \beta_k^{-1/2} - 1. 
            Re-arranging, yields: 
                \beta_k &\le 
                    1 + \frac{\alpha_0\sqrt{L_0}}{2}\sum_{i = 1}^{k} \sqrt{L_i^{-1}}
                \right)^{-2}
                    1 + \frac{k\alpha_0\sqrt{L_0}}{2\sqrt{L_{\max}}}
                \right)^{-2}. 
            Recall that $L_k = B_k + \rho_k$, and by our assumptions, the sequence $(B_k)_{k \ge 0}, (\rho_k)_{k\ge 0}$ are both bounded above, then it's not hard to see that $\beta_k \le \mathcal O(1/k^2)$. 
            Combine it with Lemma \ref{lemma:inxt-apg-cnvg-prep1}, using $\alpha_0 = 1$ will yield the desired result. 
            Finally, one can solves for $\alpha_k$ by the recursive equality, and it should yield:

\section{Linear convergence for the proximal problem}

@@

    To attain $x_k$, one approach is to utilize Theorem \ref{thm:dlty-gap-inxt-pp} numerically, usually using iterative algorithms. 

@@

            \item let $\tau > 0$ be the step size inverse, let $T_{\tau}$ be the proximal operator of $f + g$ as given by $T_{\tau}(x) := \hprox_{\tau^{-1}g}(x - \tau^{-1}\nabla f(x))$, 

@@

        Let the gradient mapping $\mathcal G_{\tau}$ be defined as: $\mathcal G_\tau(x) := \tau (x - T_\tau(x))$     

@@

            \item Let function $\Psi_\lambda$ as given by \eqref{eqn:dual-pp} which satisfies gradient mapping error bound (Assumption \ref{ass:pg-eb}) where, $f(v) = \frac{1}{2\lambda}\Vert \lambda A^\top v - y\Vert^2, g(v) = \omega^\star(v) - \frac{1}{2\lambda}\Vert y\Vert^2$. 
            \item The primal objective $\Phi_\lambda$ is a $L_1$ Lipschitz continuous on its domain. 
            \item Let $(f, g, L)$ satisfies Assumption \ref{ass:for-inxt-prox}, we can do this because $f$ is quadratic and has a Lipschitz continuous gradient. 

@@

                F(\tilde x) - F(z) 
                \frac{\tau}{2}\Vert x - z\Vert^2 - \frac{\tau}{2}\Vert z - \tilde x\Vert^2

@@

                - \frac{\tau}{2}\Vert x - \tilde x\Vert^2

@@

                &=  - \frac{1}{2\tau}\Vert \mathcal G_\tau(x) \Vert^2

@@

                &\le  - \frac{1}{2\tau}\Vert \mathcal G_\tau(x)\Vert^2 

@@

@@

            Now, for all $z = \bar x \in S$, from Assumption \ref{ass:eb-for-pp} $\exists \gamma > 0$ such that: 

@@

    \subsection{characterizing linear convergence of the proximal problem}
        In this section, we will focus on the sufficient characterization of the proximal problem which allows proximal gradient method to achieve linear convergence rate. 
        We can immediately claim quadratic growth condition when $\omega = \Vert \cdot\Vert_1$. 
        The next proposition will characterize a precise case where Assumption \ref{ass:eb-for-pp} is true, and it's a case widely available in applications. 
        \begin{proposition}[1-norm problem]
            Let $(g, \omega, A)$ satisfies Assumption \ref{ass:for-inxt-prox}. 
            In addition, if $g := \Vert \cdot\Vert_1$, then the function $\Psi_\lambda$ satisfies Assumption \ref{ass:eb-for-pp}. 
        \end{proposition}
        \begin{proof}
            Take note that the dual has closed form $g^\star(z) =\delta(z | \{x : \Vert x\Vert_1 \le 1\})$. 
            The $g^\star$ is an indicator function which represents a box constraint. 
            The objective function $\Psi_\lambda(v) = \frac{1}{2\lambda}\Vert \lambda A^\top v - y\Vert^2 + \omega^\star(v) - \frac{1}{2\lambda}\Vert y\Vert^2$ so it's Lipschitz smooth, and it must have a set of minimizers $S$ because its domain is compact. 
            In addition, $\Psi_\lambda$ fits the assumption of Necoara et al. \cite[Theorem 8]{necoara_linear_2019}. 
            Therefore, $\Psi_\lambda$ is quasi-strongly convex then by \cite[Theorem 4]{necoara_linear_2019}, and \cite[Theorem 7]{necoara_linear_2019}, it satisfies error bound condition as given in Assumption \ref{ass:pg-eb}, hence Assumption \ref{ass:eb-for-pp} also. 
            \par
            Let $\mu_f = \lambda\Vert A\Vert^2/\kappa_f$, let $\kappa_f = \theta^{-2}(A, C)$ be the Hoffman Constant as presented by Necoara et al. in \cite[Section 4]{necoara_linear_2019} where $C$ is the constraint matrix of the inequality set $\delta(z | \{x : \Vert x\Vert_1 \le 1\})$, then the error bound constant can be satisfied with
            \begin{align*}
                \gamma = \frac{\kappa_f}{1 + \mu_f + \sqrt{1 + \mu_f}}. 
            \end{align*}

        \end{proof}
        The following propositions precisely show that the linear convergence is achievable for the inner loop when $g = \Vert \cdot\Vert_1$ for our optimization objective. 
        \begin{proposition}[inner loop linear convergence scenario 1]\label{prop:inn-loop-lin-cnvg}

        \end{proposition}
    \subsection{inner loop complexity}

\section{Total complexity of the algorithm}
    This section puts results regarding the total complexity of the proposed inexact proximal gradient algorithm.
